---
title: "Tables and Graphs"
author: "Rick Morales"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r}
library(haven)
library(gtsummary)
library(dplyr)
library(ggplot2)
library(ggcorrplot)

brfss_2017 <- read_xpt("LLCP2017.XPT")

head(brfss_2017)

colnames(brfss_2017)
```


# Table 1
```{r}
brfss_2017 <- brfss_2017 %>%
  filter(!is.na(`_ECIGSTS`), ASTHMA3 %in% c(1, 2), SEX %in% c(1, 2), EDUCA <= 6, INCOME2 <= 8) %>%
  mutate(
    VapeStatus = factor(`_ECIGSTS`,
                        levels = c(1, 2, 3),
                        labels = c("Current User", "Former User", "Never Used")),
    ASTHMA3 = factor(ASTHMA3, levels = c(1, 2), labels = c("Yes", "No")),
    SEX = factor(SEX, levels = c(1, 2), labels = c("Male", "Female")),
    EducationGroup = case_when(
      EDUCA %in% c(1, 2, 3) ~ "Below High School",
      EDUCA == 4 ~ "High School Grad/GED",
      EDUCA == 5 ~ "Some College",
      EDUCA == 6 ~ "College Graduate",
      TRUE ~ NA_character_
    ),
    INCOME2 = factor(INCOME2, levels = 1:8, labels = c(
      "<$10k", "$10–15k", "$15–20k", "$20–25k", "$25–35k", "$35–50k", "$50–75k", ">$75k"
    ))
  )

table1 <- brfss_2017 %>%
  select(VapeStatus, ASTHMA3, SEX, EducationGroup, INCOME2) %>%
  tbl_summary(by = VapeStatus, missing = "ifany") %>%
  add_overall() %>%
  bold_labels()

table1
```


# Table 2
```{r}
brfss_2017_cont <- brfss_2017 %>%
  filter(!is.na(`_ECIGSTS`), !is.na(`_AGE80`), ALCDAY5 < 777) %>%
  mutate(
    VapeStatus = factor(`_ECIGSTS`,
                        levels = c(1, 2, 3),
                        labels = c("Current User", "Former User", "Never Used")),
    DRINK_DAYS = ifelse(ALCDAY5 >= 101 & ALCDAY5 <= 130, ALCDAY5 - 100,
                        ifelse(ALCDAY5 == 888, 0, NA))
  )

table2 <- brfss_2017_cont %>%
  select(VapeStatus, `_AGE80`, DRINK_DAYS) %>%
  tbl_summary(
    by = VapeStatus,
    statistic = all_continuous() ~ "{mean} (± {sd})",
    digits = all_continuous() ~ 1
  ) %>%
  add_overall() %>%
  bold_labels()

table2
```


# Graph 1
```{r}
graph1_data <- brfss_2017 %>%
  filter(`_ECIGSTS` %in% c(1, 2, 3), ASTHMA3 %in% c("Yes", "No")) %>%
  mutate(
    VapeStatus = factor(`_ECIGSTS`,
      levels = c(1, 2, 3),
      labels = c("Current User", "Former User", "Never Used"))
  )

ggplot(graph1_data, aes(x = VapeStatus, fill = ASTHMA3)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("Yes" = "red", "No" = "blue")) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "Asthma Status by Vaping Status",
    x = "Vaping Status",
    y = "Proportion",
    fill = "Asthma"
  ) +
  theme_minimal()
```


# Graph 2
```{r}
graph2_data <- brfss_2017 %>%
  filter(`_ECIGSTS` %in% c(1, 2, 3), ALCDAY5 < 888) %>%
  mutate(
    VapeStatus = factor(`_ECIGSTS`,
      levels = c(1, 2, 3),
      labels = c("Current User", "Former User", "Never Used")),
    DrinkDays = case_when(
      ALCDAY5 >= 101 & ALCDAY5 <= 130 ~ ALCDAY5 - 100,
      TRUE ~ NA_real_
    )
  ) %>%
  filter(!is.na(DrinkDays))

ggplot(graph2_data, aes(x = VapeStatus, y = DrinkDays, fill = VapeStatus)) +
  geom_boxplot() +
  scale_fill_manual(values = c(
    "Current User" = "red",
    "Former User" = "blue",
    "Never Used" = "green"
  )) +
  labs(
    title = "Alcohol Use (Days Drank in Past 30) by Vaping Status",
    x = "Vaping Status",
    y = "Drinking Days (Past 30)",
    fill = "Vaping Status"
  ) +
  theme_minimal()
```


# Graph 3
```{r}
graph3_data <- brfss_2017 %>%
  filter(`_ECIGSTS` %in% c(1, 2, 3), !is.na(`_AGE80`)) %>%
  mutate(
    VapeStatus = factor(`_ECIGSTS`,
      levels = c(1, 2, 3),
      labels = c("Current User", "Former User", "Never Used"))
  )

ggplot(graph3_data, aes(x = VapeStatus, y = `_AGE80`, fill = VapeStatus)) +
  geom_boxplot() +
  scale_fill_manual(values = c(
    "Current User" = "red",
    "Former User" = "blue",
    "Never Used" = "green"
  )) +
  labs(
    title = "Age by Vaping Status",
    x = "Vaping Status",
    y = "Age (Capped at 80+)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```


# Graph 4
```{r}
brfss_2017 <- brfss_2017 %>%
  mutate(
    EducationGroup = case_when(
      EDUCA %in% c(1, 2, 3) ~ "Below High School",
      EDUCA == 4 ~ "High School Grad/GED",
      EDUCA == 5 ~ "Some College",
      EDUCA == 6 ~ "College Graduate",
      TRUE ~ NA_character_
    ),
    EducationGroup = factor(EducationGroup,
      levels = c("Below High School", "High School Grad/GED", "Some College", "College Graduate"))
  ) %>%
  filter(!is.na(EducationGroup), !is.na(VapeStatus))

education_colors <- c(
  "Below High School" = "red",
  "High School Grad/GED" = "blue",
  "Some College" = "green",
  "College Graduate" = "yellow"
)

ggplot(brfss_2017, aes(x = VapeStatus, fill = EducationGroup)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = education_colors) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Education Level by Vaping Status",
    x = "Vaping Status",
    y = "Proportion",
    fill = "Education Level"
  ) +
  theme_minimal()
```


# Graph 5
```{r}
brfss_2017_clean <- brfss_2017 %>%
  filter(!is.na(GENHLTH), !is.na(VapeStatus)) %>%
  mutate(
    HealthStatus = factor(GENHLTH,
      levels = c(1, 2, 3, 4, 5),
      labels = c("Excellent", "Very Good", "Good", "Fair", "Poor"))
  )

health_colors <- c(
  "Excellent" = "red",
  "Very Good" = "blue",
  "Good" = "green",
  "Fair" = "yellow",
  "Poor" = "orange"
)

ggplot(brfss_2017_clean, aes(x = HealthStatus, fill = VapeStatus)) +
  geom_bar(position = "dodge") +
  labs(
    title = "General Health Status by Vaping Status",
    x = "Self-Reported General Health",
    y = "Count",
    fill = "Vaping Status"
  ) +
  scale_fill_manual(values = c("Current User" = "red", "Former User" = "blue", "Never Used" = "green")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


# Graph 6
```{r}
brfss_2017 <- brfss_2017 %>%
  mutate(
    IncomeLevel = factor(INCOME2,
      levels = c("<$10k", "$10–15k", "$15–20k", "$20–25k", "$25–35k", "$35–50k", "$50–75k", ">$75k"))
  )

income_plot_data <- brfss_2017 %>%
  filter(!is.na(VapeStatus), !is.na(IncomeLevel))

ggplot(income_plot_data, aes(x = IncomeLevel, fill = VapeStatus)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Income Distribution by Vaping Status",
    x = "Income Level",
    y = "Count",
    fill = "Vaping Status"
  ) +
  scale_fill_manual(values = c("Current User" = "red", "Former User" = "blue", "Never Used" = "green")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


# Graph 7
```{r}
brfss_2017 <- brfss_2017 %>%
  mutate(
    COPD = factor(CHCCOPD1, levels = c(1, 2), labels = c("Yes", "No"))
  ) %>%
  filter(!is.na(COPD), !is.na(VapeStatus))

ggplot(brfss_2017, aes(x = VapeStatus, fill = COPD)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("Yes" = "red", "No" = "blue")) +
  labs(
    title = "COPD Status by Vaping Status",
    x = "Vaping Status",
    y = "Proportion",
    fill = "COPD Status"
  ) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal()
```
# Correlation Matrix
```{r}
selected_vars <- brfss_2017 %>%
  select(`_AGE80`, ALCDAY5, AVEDRNK2, MAXDRNKS, `_BMI5`, `_STATE`) %>%
  filter(complete.cases(.))

cor_matrix_small <- cor(selected_vars, use = "complete.obs")

ggcorrplot(cor_matrix_small,
           method = "circle",
           type = "lower",
           lab = TRUE,
           lab_size = 4,
           title = "Correlation Matrix of Key Numeric Variables",
           show.legend = TRUE)
```
# Graph 8
Vaping Status among those who have been diagnosed with asthma in all 50 US states and territories of Guman and Puerto Rico.
```{r}
#Converting numeric variable _STATE into character and recoding with abbreviations for US states and territories
brfss_2017$`_STATE` <- as.character(brfss_2017$`_STATE`)
brfss_2017$`_STATE` <- recode(brfss_2017$'_STATE',
                       '1' = 'AL',
                       '2' = 'AK',
                       '4' = 'AZ',
                       '5' = 'AR',
                       '6' = 'CA',
                       '8' = 'CO',
                       '9' = 'CT',
                       '10' = 'DE',
                       '11' = 'DC',
                       '12' = 'FL',
                       '13' = 'GA',
                       '15' = 'HI',
                       '16' = 'ID',
                       '17' = 'IL',
                       '18' = 'IN',
                       '19' = 'IA',
                       '20' = 'KS',
                       '21' = 'KY',
                       '22' = 'LA',
                       '23' = 'ME',
                       '24' = 'MD',
                       '25' = 'MA',
                       '26' = 'MI',
                       '27' = 'MN',
                       '28' = 'MS',
                       '29' = 'MO',
                       '30' = 'MT',
                       '31' = 'NE',
                       '32' = 'NV',
                       '33' = 'NH',
                       '34' = 'NJ',
                       '35' = 'NM',
                       '36' = 'NY',
                       '37' = 'NC',
                       '38' = 'ND',
                       '39' = 'OH',
                       '40' = 'OK',
                       '41' = 'OR',
                       '42' = 'PA',
                       '44' = 'RI',
                       '45' = 'SC',
                       '46' = 'SD',
                       '47' = 'TN',
                       '48' = 'TX',
                       '49' = 'UT',
                       '50' = 'VT',
                       '51' = 'VA',
                       '53' = 'WA',
                       '54' = 'WV',
                       '55' = 'WI',
                       '56' = 'WY',
                       '66' = 'GU',
                       '72' = 'PR')

state_plot_data <- brfss_2017 %>%
  filter(!is.na(VapeStatus), !is.na(`_STATE`), !is.na(ASTHMA3), ASTHMA3 == "Yes")

state_plot_data

ggplot(state_plot_data, aes(x = `_STATE`, fill = VapeStatus)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Vaping Status in Patients Diagnosed with Asthma Across US States and Territories",
    x = "State",
    y = "Count",
    fill = "Vaping Status"
  ) +
  scale_fill_manual(values = c("Current User" = "red", "Former User" = "blue", "Never Used" = "green")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))
```
# Graoh 9
Asthma Status Among Those who Vape in all 50 US states and territories of Guam and Puerto Rico. This refers to patients survyed who have been diagnosed with asthma and those who did not receive a diagnosis.
```{r}
state_plot_data2 <- brfss_2017 %>%
  filter(!is.na(ASTHMA3), !is.na(`_STATE`), !is.na(VapeStatus), VapeStatus == "Current User")
    #group_by(`_STATE`) %>%
    #summarize(coumt_current_user = n())

state_plot_data2

ggplot(state_plot_data2, aes(x = `_STATE`, fill = ASTHMA3)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Asthma Status in Current Vapers Across US States and Territories",
    x = "State",
    y = "Count",
    fill = "Asthma Status"
  ) +
  scale_fill_manual(values = c("Yes" = "red", "No" = "blue")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))
```
Graoh 10
Vaping Status among those surveyed in 2017 who report that they still have asthma across all 50 US states and territories including Guam and Puerto Rico.
```{r}
brfss_2017 <- brfss_2017 %>%
  mutate(
    ASTHNOW = factor(ASTHNOW,
      levels = c(1, 2), labels = c("Yes", "No"))
  )

asthma_now_plot_data <- brfss_2017 %>%
  filter(!is.na(VapeStatus), !is.na(`_STATE`), !is.na(ASTHNOW), ASTHNOW == "Yes")

ggplot(asthma_now_plot_data, aes(x = `_STATE`, fill = VapeStatus)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Vaping Status in Current Asthmatics Across by State/Territory",
    x = "State",
    y = "Count",
    fill = "Vaping Status"
  ) +
  scale_fill_manual(values = c("Current User" = "red", "Former User" = "blue", "Never Used" = "green")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))
```
Sex Distribution by Vaping Status
```{r}
brfss_2017 <- brfss_2017 %>%
  filter(!is.na(SEX), !is.na(VapeStatus))

ggplot(brfss_2017, aes(x = VapeStatus, fill = SEX)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("Male" = "red", "Female" = "blue")) +
  labs(
    title = "Sex by Vaping Status",
    x = "Vaping Status",
    y = "Proportion",
    fill = "Sex"
  ) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal()
```


# Missing Data Summary Table
```{r}
library(naniar)

missing_summary <- brfss_2017 %>%
  select(ASTHMA3, SEX, EDUCA, INCOME2, GENHLTH, `_AGE80`, ALCDAY5, AVEDRNK2, MAXDRNKS, `_BMI5`) %>%
  miss_var_summary()

print(missing_summary)
```


