---
title: "IM2"
author: "Rick Morales"
date: "`r Sys.Date()`"
output: pdf_document
---

# 1. Libraries
```{r}
library(haven)
library(dplyr)
library(ggplot2)
library(glmnet)
library(parameters)
library(see)
library(car)
library(caret)
library(pROC)
library(ggplot2)
library(randomForest)
library(mice)
library(tidyr)
```

# 2. Load Data
```{r}
brfss_2017 <- read_xpt("LLCP2017.XPT")
```


# 3. Clean Data (carry over from Preprocessing)
```{r}
brfss_2017_cleaned <- brfss_2017 %>%
  select(-c(2:6), -c(10:27), -31, -37, -40, -45, -c(53:57), -c(62:64), -72,
         -c(80:82), -85, -c(87:89), -c(97:102), -106, -108, -110, -c(112:246),
         -c(255:258), -c(326:327), -c(330:337), -c(340:345), -c(356:357))
```


# 4.Renamed For Clarity
```{r}
names(brfss_2017_cleaned)[1] <- "STATE"
names(brfss_2017_cleaned)[4] <- "PSU"
names(brfss_2017_cleaned)[8] <- "HLTHPLN"
names(brfss_2017_cleaned)[9] <- "PERSDOC"
names(brfss_2017_cleaned)[11] <- "CHECKUP"
names(brfss_2017_cleaned)[12] <- "BPHIGH"
names(brfss_2017_cleaned)[13] <- "CHOLCHK"
names(brfss_2017_cleaned)[14] <- "TOLDHI"
names(brfss_2017_cleaned)[15] <- "CVDINFR"
names(brfss_2017_cleaned)[16] <- "CVDCRHD"
names(brfss_2017_cleaned)[17] <- "CVDSTRK"
names(brfss_2017_cleaned)[18] <- "ASTHMA"
names(brfss_2017_cleaned)[21] <- "CHCCOPD"
names(brfss_2017_cleaned)[22] <- "HAVARTH"
names(brfss_2017_cleaned)[23] <- "ADDEPEV"
names(brfss_2017_cleaned)[25] <- "DIABETES"
names(brfss_2017_cleaned)[29] <- "RENTHOME"
names(brfss_2017_cleaned)[30] <- "VETERAN"
names(brfss_2017_cleaned)[31] <- "EMPLOYED"
names(brfss_2017_cleaned)[33] <- "INCOME"
names(brfss_2017_cleaned)[35] <- "WEIGHT"
names(brfss_2017_cleaned)[36] <- "HEIGHT"
names(brfss_2017_cleaned)[44] <- "USENOW"
names(brfss_2017_cleaned)[47] <- "FRUIT"
names(brfss_2017_cleaned)[48] <- "FRUIT_JUICE"
names(brfss_2017_cleaned)[49] <- "GREEN_VEG"
names(brfss_2017_cleaned)[50] <- "FRENCHF"
names(brfss_2017_cleaned)[51] <- "POTATO"
names(brfss_2017_cleaned)[52] <- "OTHER_VEG"
names(brfss_2017_cleaned)[53] <- "EXERCISE_ANY"
names(brfss_2017_cleaned)[56] <- "FLUSHOT"
names(brfss_2017_cleaned)[57] <- "PNEUMONIA_VAC"
names(brfss_2017_cleaned)[58] <- "HIVTEST"
names(brfss_2017_cleaned)[59] <- "HIVRISK"
names(brfss_2017_cleaned)[63] <- "STSTR"
names(brfss_2017_cleaned)[64] <- "STRATUM_WGHT"
names(brfss_2017_cleaned)[65] <- "RAWRAKE"
names(brfss_2017_cleaned)[66] <- "WT2RAKE"
names(brfss_2017_cleaned)[67] <- "IMPRACE"
names(brfss_2017_cleaned)[68] <- "DUALUSE"
names(brfss_2017_cleaned)[69] <- "DUALCOR"
names(brfss_2017_cleaned)[70] <- "LLCPWT2"
names(brfss_2017_cleaned)[71] <- "LLCPWT"
names(brfss_2017_cleaned)[72] <- "RFHLTH"
names(brfss_2017_cleaned)[73] <- "PHYS_COMP"
names(brfss_2017_cleaned)[74] <- "MENT_COMP"
names(brfss_2017_cleaned)[75] <- "HCVU651"
names(brfss_2017_cleaned)[76] <- "RFHYPES"
names(brfss_2017_cleaned)[77] <- "CHOLCH1"
names(brfss_2017_cleaned)[78] <- "HICHOL_COMP"
names(brfss_2017_cleaned)[79] <- "MI_CHD"
names(brfss_2017_cleaned)[80] <- "ASTHMA_LIFE"
names(brfss_2017_cleaned)[81] <- "ASTHMA_NOW_COMP"
names(brfss_2017_cleaned)[82] <- "ASTHMA_COMP"
names(brfss_2017_cleaned)[83] <- "ARTHRITIS_DIAGNOS"
names(brfss_2017_cleaned)[84] <- "LIMIT_ACTIVITY"
names(brfss_2017_cleaned)[85] <- "LIMIT_WORK"
names(brfss_2017_cleaned)[86] <- "LIMIT_SOCIAL"
names(brfss_2017_cleaned)[87] <- "PRACE_COMP"
names(brfss_2017_cleaned)[88] <- "MULTI_RACE"
names(brfss_2017_cleaned)[89] <- "HISPANIC_COMP"
names(brfss_2017_cleaned)[90] <- "RACE_COMP"
names(brfss_2017_cleaned)[91] <- "RACEG21"
names(brfss_2017_cleaned)[92] <- "RACEGR3"
names(brfss_2017_cleaned)[93] <- "RACE_G1"
names(brfss_2017_cleaned)[94] <- "AGEG5YR"
names(brfss_2017_cleaned)[95] <- "AGE65YR"
names(brfss_2017_cleaned)[96] <- "AGE80"
names(brfss_2017_cleaned)[97] <- "AGE_G"
names(brfss_2017_cleaned)[98] <- "HT_COMP_IN"
names(brfss_2017_cleaned)[99] <- "HT_COMP_M"
names(brfss_2017_cleaned)[100] <- "WT_COMP_KILO"
names(brfss_2017_cleaned)[101] <- "BMI"
names(brfss_2017_cleaned)[102] <- "BMI_CAT"
names(brfss_2017_cleaned)[103] <- "BMI_OVER"
names(brfss_2017_cleaned)[104] <- "CHILD_COMP"
names(brfss_2017_cleaned)[105] <- "EDUCA_COMP"
names(brfss_2017_cleaned)[106] <- "INCOME_COMP"
names(brfss_2017_cleaned)[107] <- "SMOKER_COMP"
names(brfss_2017_cleaned)[108] <- "SMOKENOW_COMP"
names(brfss_2017_cleaned)[109] <- "ECIG_COMP"
names(brfss_2017_cleaned)[110] <- "ECIG_NOW_COMP"
names(brfss_2017_cleaned)[111] <- "DRINKANY_30"
names(brfss_2017_cleaned)[112] <- "DROCDY3"
names(brfss_2017_cleaned)[113] <- "BINGE_COMP"
names(brfss_2017_cleaned)[114] <- "WEEKLY_DRINK"
names(brfss_2017_cleaned)[115] <- "HEAVY_DRINK"
names(brfss_2017_cleaned)[116] <- "FRUITJUICE_COMP"
names(brfss_2017_cleaned)[117] <- "FRUITJUICE_COMP2"
names(brfss_2017_cleaned)[118] <- "GREEN_VEG_COMP"
names(brfss_2017_cleaned)[119] <- "FRENCHF_COMP"
names(brfss_2017_cleaned)[120] <- "POTATO_COMP"
names(brfss_2017_cleaned)[121] <- "OTHER_VEG_COMP"
names(brfss_2017_cleaned)[122] <- "MISSING_FRUIT"
names(brfss_2017_cleaned)[123] <- "MISSING_VEG"
names(brfss_2017_cleaned)[124] <- "MISS_ANY_FRUIT"
names(brfss_2017_cleaned)[125] <- "MISS_ANY_VEG"
names(brfss_2017_cleaned)[126] <- "TOTAL_FRUIT"
names(brfss_2017_cleaned)[127] <- "TOTAL_VEG"
names(brfss_2017_cleaned)[128] <- "ATLEAST1_FRUIT"
names(brfss_2017_cleaned)[129] <- "ATLEAST1_VEG"
names(brfss_2017_cleaned)[130] <- "MORE16_FRUIT"
names(brfss_2017_cleaned)[131] <- "MORE23_VEG"
names(brfss_2017_cleaned)[132] <- "FRUIT_EXCLUDE"
names(brfss_2017_cleaned)[133] <- "VEG_EXCLUDE"
names(brfss_2017_cleaned)[134] <- "LEISURE_COMP"
names(brfss_2017_cleaned)[135] <- "MAX_OXYGEN"
names(brfss_2017_cleaned)[136] <- "FUNC_CAPACITY"
names(brfss_2017_cleaned)[137] <- "STRENGTH_FREQ"
names(brfss_2017_cleaned)[138] <- "MISSING_PHYS"
names(brfss_2017_cleaned)[139] <- "PHYS_CATEGORY"
names(brfss_2017_cleaned)[140] <- "PHYS_IDX"
names(brfss_2017_cleaned)[141] <- "150_PHYS_COMP"
names(brfss_2017_cleaned)[142] <- "300_PHYS_COMP"
names(brfss_2017_cleaned)[143] <- "300_PHYS2_COMP"
names(brfss_2017_cleaned)[144] <- "MUSCLE_RECOMMEND"
names(brfss_2017_cleaned)[145] <- "AEROBIC_STRENGTH"
names(brfss_2017_cleaned)[146] <- "AEROBIC_STRENGTH2"
names(brfss_2017_cleaned)[147] <- "SEATBELT_COMP"
names(brfss_2017_cleaned)[148] <- "SEATBELT_COMP2"
names(brfss_2017_cleaned)[149] <- "AIDSTEST_COMP"


head(brfss_2017_cleaned)
```


# 5. Missing Value Summary
```{r}
# Summarize missing values (count and %)
missing_summary <- brfss_2017_cleaned %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Missing_Count") %>%
  mutate(Missing_Percent = round((Missing_Count / nrow(brfss_2017_cleaned)) * 100, 2)) %>%
  arrange(desc(Missing_Count)) %>%
  filter(Missing_Count > 0)

# Show missing value table
print(missing_summary)

# Optional: Visualize top 15 missing variables
top_missing <- missing_summary %>% slice_max(order_by = Missing_Count, n = 25)

ggplot(top_missing, aes(x = reorder(Variable, Missing_Count), y = Missing_Count)) +
  geom_col(fill = "tomato") +
  coord_flip() +
  labs(title = "Top 25 Variables with Missing Values",
       x = "Variable",
       y = "Missing Count") +
  theme_minimal()
```


# 6. Prepare Dataset (filtered and relevant variables only) and Imputation
```{r}
# Ensure unique column names
colnames(brfss_2017_cleaned) <- make.names(colnames(brfss_2017_cleaned), unique = TRUE)

# Replace NAs
brfss_2017_cleaned[is.na(brfss_2017_cleaned)] <- 0

# Filter for ECIG_COMP values and asthmatic respondents
df <- brfss_2017_cleaned %>%
  filter(ECIG_COMP != 0 & ASTHMA_NOW_COMP == 1)

# Compute BMI from height and weight, then standardize it
# height in meters, weight in kg
# Height in inches is variable HEIGHT, Weight in lbs is WEIGHT
# BMI = (weight in pounds / (height in inches)^2) * 703

df <- df %>%
  mutate(BMI_COMP = (WEIGHT / (HEIGHT^2)) * 703)

# Split into Training and Test sets
set.seed(123)
train_indices <- sample(seq_len(nrow(df)), size = 0.8 * nrow(df))
train_data <- df[train_indices, ]
test_data <- df[-train_indices, ]

# MICE Imputation on Training Data
# Check missing summary first
md.pattern(train_data)

# Apply MICE with CART method (Classification and Regression Trees)
mice_train <- mice(train_data, method = "cart", m = 1, seed = 123)

# Complete the data (first imputed dataset)
train_data_imputed <- complete(mice_train)

# Optional: Check again
summary(train_data_imputed)
```


# 7. Predictor and Target Setup
```{r}
predictors <- c("AGE80", "BMI_COMP", "SMOKER_COMP", "STATE", "ADDEPEV",
                "CHCCOPD", "CHECKUP", "GENHLTH", "HLTHPLN",
                "EDUCA_COMP", "INCOME_COMP", "SEX", "WEEKLY_DRINK")

x <- data.matrix(df[, predictors])
y <- df$ECIG_COMP

data <- data.frame(x, y)
```


# 8. Linear Model
```{r}
model <- lm(y ~ ., data = data)
summary(model)
```


# 9. Model Assumptions
```{r}
# Visualize residuals
plot(model)

# Check model parameters
model_parameters(model)

# VIF (multicollinearity)
vif(model)
```


# 10. Cross-Validation (Lasso for example)
```{r}
set.seed(123)
cv_model <- train(
  x = x,
  y = y,
  method = "glmnet",
  trControl = trainControl(method = "cv", number = 5)
)
print(cv_model)
```


# 11. RMSE Calculation
```{r}
predictions <- predict(model, newdata = data.frame(x))
actual_values <- y
rmse <- sqrt(mean((actual_values - predictions)^2))
print(paste("Train RMSE:", round(rmse, 3)))
```


# 12. Visualization
```{r}
# Residual Plot
ggplot(data.frame(resid = residuals(model)), aes(x = resid)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  labs(title = "Residuals Distribution", x = "Residuals")

# Coefficient Importance Plot
coef_df <- data.frame(Variable = names(coef(model))[-1], Coefficient = coef(model)[-1])
ggplot(coef_df, aes(x = reorder(Variable, abs(Coefficient)), y = Coefficient)) +
  geom_col(fill = "orange") +
  coord_flip() +
  labs(title = "Coefficient Estimates", x = "Variables", y = "Coefficient")
```


# 13. Logistic Model (Binary: Current User vs. Others)
```{r}
df$ECIG_BINARY <- ifelse(df$ECIG_COMP == 1 | df$ECIG_COMP == 2, 1, 0)
log_model <- glm(ECIG_BINARY ~ ., data = data.frame(df[, predictors], ECIG_BINARY = df$ECIG_BINARY), family = "binomial")
summary(log_model)
```


# 14. ROC and AUC Plot
```{r}
probs <- predict(log_model, type = "response")
roc_obj <- roc(df$ECIG_BINARY, probs)
auc(roc_obj)

plot(roc_obj, col = "blue", lwd = 2, main = "ROC Curve")
abline(a = 0, b = 1, lty = 2, col = "red")
```


# 15. Calibration Plot
```{r}
# Recalculate probabilities (if needed)
probs <- predict(log_model, type = "response")

# Create a dataframe for plotting
calib_data <- data.frame(
  predicted = probs,
  actual = factor(df$ECIG_BINARY)
)

# Use caret to create calibration data
calib <- calibration(actual ~ predicted, data = calib_data, class = "1", cuts = 10)

# Plot the calibration curve
ggplot(calib) +
  geom_line(aes(x = midpoint, y = Percent), color = "blue", linewidth = 1.2) +
  geom_point(aes(x = midpoint, y = Percent), color = "blue", size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", color = "gray40") +
  labs(title = "Calibration Plot",
       x = "Predicted Probability",
       y = "Observed Proportion (Current User)") +
  theme_minimal(base_size = 14)
```


# 16. Random Forest Model + ROC/AUC
```{r}
# Convert outcome to factor for classification
df$ECIG_BINARY <- factor(ifelse(df$ECIG_COMP == 1 | df$ECIG_COMP == 2, 1, 0))

# Train Random Forest model
set.seed(123)
rf_model <- randomForest(
  x = df[, predictors],
  y = df$ECIG_BINARY,
  ntree = 100,           # number of trees
  mtry = 3,              # number of variables per split (can tune)
  importance = TRUE
)

# Check model summary
print(rf_model)
```


# 17. Random Forest ROC & AUC
```{r}
# Predicted probabilities
rf_probs <- predict(rf_model, type = "prob")[, 2]  # probability of class "1"

# ROC curve
rf_roc <- roc(df$ECIG_BINARY, rf_probs)

# AUC
auc(rf_roc)

# Plot ROC
plot(rf_roc, col = "forestgreen", lwd = 2, main = "Random Forest ROC Curve")
abline(a = 0, b = 1, lty = 2, col = "gray")
```


# 18. Feature Importance
```{r}
# Visualize feature importance
varImpPlot(rf_model, main = "Variable Importance (Random Forest)", pch = 19, col = "darkgreen")
```






