---
title: "Initial Model"
author: "Rick Morales"
date: "`r Sys.Date()`"
output: pdf_document
---

# Libraries
```{r}
library(haven)
library(mice) # For imputation
library(tidyverse)
library(caret)
library(glmnet)
library(parameters)
library(see)
library(lmtest)  # For assumption testing
library(ggplot2)
library(car)  # For Variance Inflation Factor (VIF)

brfss_2017 <- read_xpt("LLCP2017.XPT")

head(brfss_2017)

colnames(brfss_2017)
```


# Cleaned Data
```{r}
brfss_2017_cleaned <- brfss_2017 %>%
    select(-c(2:6), -c(10:27), -31, -37, -40, -45, -c(53:57), -c(62:64), -72, -c(80:82), -85, -c(87:89), -c(97:102), -106, -108, -110, -c(112:246), ... = -c(255:258), -c(326:327), -c(330:337), -c(340:345), -c(356:357))

head(brfss_2017_cleaned)
```


# Renamed For Clarity
```{r}
names(brfss_2017_cleaned)[1] <- "STATE"
names(brfss_2017_cleaned)[4] <- "PSU"
names(brfss_2017_cleaned)[8] <- "HLTHPLN"
names(brfss_2017_cleaned)[9] <- "PERSDOC"
names(brfss_2017_cleaned)[11] <- "CHECKUP"
names(brfss_2017_cleaned)[12] <- "BPHIGH"
names(brfss_2017_cleaned)[13] <- "CHOLCHK"
names(brfss_2017_cleaned)[14] <- "TOLDHI"
names(brfss_2017_cleaned)[15] <- "CVDINFR"
names(brfss_2017_cleaned)[16] <- "CVDCRHD"
names(brfss_2017_cleaned)[17] <- "CVDSTRK"
names(brfss_2017_cleaned)[18] <- "ASTHMA"
names(brfss_2017_cleaned)[21] <- "CHCCOPD"
names(brfss_2017_cleaned)[22] <- "HAVARTH"
names(brfss_2017_cleaned)[23] <- "ADDEPEV"
names(brfss_2017_cleaned)[25] <- "DIABETES"
names(brfss_2017_cleaned)[29] <- "RENTHOME"
names(brfss_2017_cleaned)[30] <- "VETERAN"
names(brfss_2017_cleaned)[31] <- "EMPLOYED"
names(brfss_2017_cleaned)[33] <- "INCOME"
names(brfss_2017_cleaned)[35] <- "WEIGHT"
names(brfss_2017_cleaned)[36] <- "HEIGHT"
names(brfss_2017_cleaned)[44] <- "USENOW"
names(brfss_2017_cleaned)[47] <- "FRUIT"
names(brfss_2017_cleaned)[48] <- "FRUIT_JUICE"
names(brfss_2017_cleaned)[49] <- "GREEN_VEG"
names(brfss_2017_cleaned)[50] <- "FRENCHF"
names(brfss_2017_cleaned)[51] <- "POTATO"
names(brfss_2017_cleaned)[52] <- "OTHER_VEG"
names(brfss_2017_cleaned)[53] <- "EXERCISE_ANY"
names(brfss_2017_cleaned)[56] <- "FLUSHOT"
names(brfss_2017_cleaned)[57] <- "PNEUMONIA_VAC"
names(brfss_2017_cleaned)[58] <- "HIVTEST"
names(brfss_2017_cleaned)[59] <- "HIVRISK"
names(brfss_2017_cleaned)[63] <- "STSTR"
names(brfss_2017_cleaned)[64] <- "STRATUM_WGHT"
names(brfss_2017_cleaned)[65] <- "RAWRAKE"
names(brfss_2017_cleaned)[66] <- "WT2RAKE"
names(brfss_2017_cleaned)[67] <- "IMPRACE"
names(brfss_2017_cleaned)[68] <- "DUALUSE"
names(brfss_2017_cleaned)[69] <- "DUALCOR"
names(brfss_2017_cleaned)[70] <- "LLCPWT2"
names(brfss_2017_cleaned)[71] <- "LLCPWT"
names(brfss_2017_cleaned)[72] <- "RFHLTH"
names(brfss_2017_cleaned)[73] <- "PHYS_COMP"
names(brfss_2017_cleaned)[74] <- "MENT_COMP"
names(brfss_2017_cleaned)[75] <- "HCVU651"
names(brfss_2017_cleaned)[76] <- "RFHYPES"
names(brfss_2017_cleaned)[77] <- "CHOLCH1"
names(brfss_2017_cleaned)[78] <- "HICHOL_COMP"
names(brfss_2017_cleaned)[79] <- "MI_CHD"
names(brfss_2017_cleaned)[80] <- "ASTHMA_LIFE"
names(brfss_2017_cleaned)[81] <- "ASTHMA_NOW_COMP"
names(brfss_2017_cleaned)[82] <- "ASTHMA_COMP"
names(brfss_2017_cleaned)[83] <- "ARTHRITIS_DIAGNOS"
names(brfss_2017_cleaned)[84] <- "LIMIT_ACTIVITY"
names(brfss_2017_cleaned)[85] <- "LIMIT_WORK"
names(brfss_2017_cleaned)[86] <- "LIMIT_SOCIAL"
names(brfss_2017_cleaned)[87] <- "PRACE_COMP"
names(brfss_2017_cleaned)[88] <- "MULTI_RACE"
names(brfss_2017_cleaned)[89] <- "HISPANIC_COMP"
names(brfss_2017_cleaned)[90] <- "RACE_COMP"
names(brfss_2017_cleaned)[91] <- "RACEG21"
names(brfss_2017_cleaned)[92] <- "RACEGR3"
names(brfss_2017_cleaned)[93] <- "RACE_G1"
names(brfss_2017_cleaned)[94] <- "AGEG5YR"
names(brfss_2017_cleaned)[95] <- "AGE65YR"
names(brfss_2017_cleaned)[96] <- "AGE80"
names(brfss_2017_cleaned)[97] <- "AGE_G"
names(brfss_2017_cleaned)[98] <- "HT_COMP_IN"
names(brfss_2017_cleaned)[99] <- "HT_COMP_M"
names(brfss_2017_cleaned)[100] <- "WT_COMP_KILO"
names(brfss_2017_cleaned)[101] <- "BMI"
names(brfss_2017_cleaned)[102] <- "BMI_CAT"
names(brfss_2017_cleaned)[103] <- "BMI_OVER"
names(brfss_2017_cleaned)[104] <- "CHILD_COMP"
names(brfss_2017_cleaned)[105] <- "EDUCA_COMP"
names(brfss_2017_cleaned)[106] <- "INCOME_COMP"
names(brfss_2017_cleaned)[107] <- "SMOKER_COMP"
names(brfss_2017_cleaned)[108] <- "SMOKENOW_COMP"
names(brfss_2017_cleaned)[109] <- "ECIG_COMP"
names(brfss_2017_cleaned)[110] <- "ECIG_NOW_COMP"
names(brfss_2017_cleaned)[111] <- "DRINKANY_30"
names(brfss_2017_cleaned)[112] <- "DROCDY3"
names(brfss_2017_cleaned)[113] <- "BINGE_COMP"
names(brfss_2017_cleaned)[114] <- "WEEKLY_DRINK"
names(brfss_2017_cleaned)[115] <- "HEAVY_DRINK"
names(brfss_2017_cleaned)[116] <- "FRUITJUICE_COMP"
names(brfss_2017_cleaned)[117] <- "FRUITJUICE_COMP2"
names(brfss_2017_cleaned)[118] <- "GREEN_VEG_COMP"
names(brfss_2017_cleaned)[119] <- "FRENCHF_COMP"
names(brfss_2017_cleaned)[120] <- "POTATO_COMP"
names(brfss_2017_cleaned)[121] <- "OTHER_VEG_COMP"
names(brfss_2017_cleaned)[122] <- "MISSING_FRUIT"
names(brfss_2017_cleaned)[123] <- "MISSING_VEG"
names(brfss_2017_cleaned)[124] <- "MISS_ANY_FRUIT"
names(brfss_2017_cleaned)[125] <- "MISS_ANY_VEG"
names(brfss_2017_cleaned)[126] <- "TOTAL_FRUIT"
names(brfss_2017_cleaned)[127] <- "TOTAL_VEG"
names(brfss_2017_cleaned)[128] <- "ATLEAST1_FRUIT"
names(brfss_2017_cleaned)[129] <- "ATLEAST1_VEG"
names(brfss_2017_cleaned)[130] <- "MORE16_FRUIT"
names(brfss_2017_cleaned)[131] <- "MORE23_VEG"
names(brfss_2017_cleaned)[132] <- "FRUIT_EXCLUDE"
names(brfss_2017_cleaned)[133] <- "VEG_EXCLUDE"
names(brfss_2017_cleaned)[134] <- "LEISURE_COMP"
names(brfss_2017_cleaned)[135] <- "MAX_OXYGEN"
names(brfss_2017_cleaned)[136] <- "FUNC_CAPACITY"
names(brfss_2017_cleaned)[137] <- "STRENGTH_FREQ"
names(brfss_2017_cleaned)[138] <- "MISSING_PHYS"
names(brfss_2017_cleaned)[139] <- "PHYS_CATEGORY"
names(brfss_2017_cleaned)[140] <- "PHYS_IDX"
names(brfss_2017_cleaned)[141] <- "150_PHYS_COMP"
names(brfss_2017_cleaned)[142] <- "300_PHYS_COMP"
names(brfss_2017_cleaned)[143] <- "300_PHYS2_COMP"
names(brfss_2017_cleaned)[144] <- "MUSCLE_RECOMMEND"
names(brfss_2017_cleaned)[145] <- "AEROBIC_STRENGTH"
names(brfss_2017_cleaned)[146] <- "AEROBIC_STRENGTH2"
names(brfss_2017_cleaned)[147] <- "SEATBELT_COMP"
names(brfss_2017_cleaned)[148] <- "SEATBELT_COMP2"
names(brfss_2017_cleaned)[149] <- "AIDSTEST_COMP"

head(brfss_2017_cleaned)
```


# Drop source variables and target leakage risks
```{r}
# Remove certain calculated variables
brfss_2017_cleaned <- brfss_2017_cleaned %>%
  select(-c(35, 36, 43, 62, 63, 64, 65, 66, 68, 69, 70, 71, 72, 73, 74, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 97, 98, 99, 100, 102, 103, 104, 105, 106, 108, 110, 111, 112, 113, 114, 115, 116, 117, 118, 119, 120, 121, 122, 123, 124, 125, 126, 127, 128, 129, 130, 131, 132, 133, 134, 135, 136, 137, 138, 139, 140, 141, 142, 143, 144, 145, 146, 147, 148, 149))

# Remove SEQNO since it's a unique identifier
brfss_2017_cleaned <- brfss_2017_cleaned %>%
    select(-c(SEQNO))

# Filter for only asthmatics
brfss_2017_cleaned <- brfss_2017_cleaned %>%
  filter(ASTHMA == 1)

# Correlation for all variables
round(cor(brfss_2017_cleaned), digits = 2)

# Create model for VIF
vif_model <- lm(ECIG_COMP ~ AGE80 + BMI + SMOKER_COMP + STATE + ADDEPEV + ALCDAY5 + CHCCOPD + CHECKUP + EDUCA + GENHLTH + HLTHPLN + INCOME + SEX, data = brfss_2017_cleaned)

# Calculating VIF
vif_values <- vif(vif_model)
vif_values

# Visualizing the model
plot(vif_model, which = 1, main = "Model Fit")

# Visualizing VIF
barplot(vif_values, las = 2, col = "skyblue", main = "Variance Inflation Factor (VIF)")

# Creating a correlation matrix
cor_matrix <- cor(brfss_2017_cleaned[c('AGE80', 'BMI', 'SMOKER_COMP', 'STATE', 'ADDEPEV', 'ALCDAY5',
                    'CHCCOPD', 'CHECKUP', 'EDUCA', 'GENHLTH', 'HLTHPLN', 'INCOME', 'SEX')])

# Visualizing the correlation matrix
image(cor_matrix, main = "Correlation Matrix", col = colorRampPalette(c("blue", "white", "red"))(20))
```


# Train-Test Split
```{r}
set.seed(123)
split <- sample(nrow(brfss_2017_cleaned), 0.6 * nrow(brfss_2017_cleaned))
train <- brfss_2017_cleaned[split, ]
test <- brfss_2017_cleaned[-split, ]
```

# Missing Value Summary
```{r}
train
missing_summary_train <- train %>%
  miss_var_summary()

print(missing_summary_train)
```

# Imputation of Missing Values

Survey data assumes missing values are classified as missing at random. Multiple imputation by chained equations. Best for survey data. Method chosen is cart (imputation by classification and regression trees)
```{r}
# Calculate imputed values with training data
methods(mice)
imp <- mice(train, method = "cart", m = 1, maxit = 1, seed = 123)
#imputed_values <- mice(train, method = "rf)
```
```{r}
complete(imp)
```
```{r}
train <- complete(imp)
train
```

# Predictor Variables
```{r}
predictor_vars <- c('AGE80', 'BMI', 'SMOKER_COMP', 'STATE', 'ADDEPEV', 'ALCDAY5',
                    'CHCCOPD', 'CHECKUP', 'EDUCA', 'GENHLTH', 'HLTHPLN', 'INCOME', 'SEX')
```


# Model Fitting (Linear Model)
```{r}
x_train <- data.matrix(train[, predictor_vars])
y_train <- train$ECIG_COMP

# Combine x_train and y_train into one data frame
train_lm <- data.frame(y_train, x_train)

# Fit the model properly using all predictors
model <- lm(y_train ~ ., data = train_lm)
summary(model)
```


# Assumption Checks
```{r}
# Residual plots
plot(model)

# Test for Autocorrelation
dwtest(model)
```


# Cross-Validation (5-Fold using caret + glmnet)
```{r}
# Remove rows with missing values in either predictors or target
complete_rows <- complete.cases(x_train, y_train)
x_train_clean <- x_train[complete_rows, ]
y_train_clean <- y_train[complete_rows]

# Then run cross-validation
cv_model <- train(x = x_train_clean, y = y_train_clean,
                  method = "glmnet",
                  trControl = trainControl(method = "cv", number = 5))

print(cv_model)
```


# Overfitting Check
```{r}
# Rebuild x_test
x_test <- test[, predictor_vars]
x_test <- data.frame(x_test)

# Predict using the linear model
predictions <- predict(model, newdata = x_test)

# Remove any rows with NA values in either predictions or actual values
valid_rows <- complete.cases(predictions, test$ECIG_COMP)
y_test_clean <- test$ECIG_COMP[valid_rows]
predictions_clean <- predictions[valid_rows]

# Calculate RMSE
rmse_train <- sqrt(mean((y_train - predict(model))^2))
rmse_test <- sqrt(mean((y_test_clean - predictions_clean)^2))

print(paste("Train RMSE:", round(rmse_train, 3)))
print(paste("Test RMSE:", round(rmse_test, 3)))
```


# RMSE Comparison
```{r}
# Define RMSE values
rmse_train <- 1.253
rmse_test <- 0.513

# Create data frame
rmse_data <- data.frame(
  Dataset = c("Train RMSE", "Test RMSE"),
  RMSE = c(rmse_train, rmse_test)
)

# Create the bar plot
ggplot(rmse_data, aes(x = Dataset, y = RMSE)) +
  geom_bar(stat = "identity", fill = "orange", width = 0.6) +
  geom_text(aes(label = round(RMSE, 3)), vjust = -0.5, size = 5) +
  labs(
    title = "Root Mean Squared Error (RMSE)",
    y = "RMSE Value",
    x = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "none"
  )
```











