---
title: "Preprocessing & Feature Engineering"
author: "Stephanie Halsing"
date: "2025-04-14"
output: html_document
---

```{r}
library(haven)
library(gtsummary)
library(dplyr)
library(ggplot2)
library(ggcorrplot)
library(naniar)
library(dplyr)
library(tidyr)
library(glmnet)
library(randomForest)
library(SPlit)
library(parameters)
library(see)
library(caret)
library(factoextra)

brfss_2017 <- read_xpt("LLCP2017.XPT")

head(brfss_2017)

colnames(brfss_2017)
```


# Pre-Processing

Additional Cleaning

Removal of missing data consists of utilizing the select function from the dplyr library. Variables with no missing values such as FMONTH and IDATE are removed due to irrelevance in the study. Columns containing 30% or more missing values are removed from the dataset, decreasing the total # of columns from 358 to 149.
```{r}
brfss_2017_cleaned <- brfss_2017 %>%
    select(-c(2:6), -c(10:27), -31, -37, -40, -45, -c(53:57), -c(62:64), -72, -c(80:82), -85, -c(87:89), -c(97:102), -106, -108, -110, -c(112:246), ... = -c(255:258), -c(326:327), -c(330:337), -c(340:345), -c(356:357))

head(brfss_2017_cleaned)
```

Variables below were renamed to improve understanding and code readability. Inserted characters such as underscores before variable names are removed. Variable names are consistent case-wise with all column names in uppercase format. Since data is already in wide format, pivoting is not needed. ECT.
```{r}
#Duplicated row check
#sum(duplicated(brfss_2017_cleaned))

names(brfss_2017_cleaned)[1] <- "STATE"
names(brfss_2017_cleaned)[4] <- "PSU"
names(brfss_2017_cleaned)[8] <- "HLTHPLN"
names(brfss_2017_cleaned)[9] <- "PERSDOC"
names(brfss_2017_cleaned)[11] <- "CHECKUP"
names(brfss_2017_cleaned)[12] <- "BPHIGH"
names(brfss_2017_cleaned)[13] <- "CHOLCHK"
names(brfss_2017_cleaned)[14] <- "TOLDHI"
names(brfss_2017_cleaned)[15] <- "CVDINFR"
names(brfss_2017_cleaned)[16] <- "CVDCRHD"
names(brfss_2017_cleaned)[17] <- "CVDSTRK"
names(brfss_2017_cleaned)[18] <- "ASTHMA"
names(brfss_2017_cleaned)[21] <- "CHCCOPD"
names(brfss_2017_cleaned)[22] <- "HAVARTH"
names(brfss_2017_cleaned)[23] <- "ADDEPEV"
names(brfss_2017_cleaned)[25] <- "DIABETES"
names(brfss_2017_cleaned)[29] <- "RENTHOME"
names(brfss_2017_cleaned)[30] <- "VETERAN"
names(brfss_2017_cleaned)[31] <- "EMPLOYED"
names(brfss_2017_cleaned)[33] <- "INCOME"
names(brfss_2017_cleaned)[35] <- "WEIGHT"
names(brfss_2017_cleaned)[36] <- "HEIGHT"
names(brfss_2017_cleaned)[44] <- "USENOW"
names(brfss_2017_cleaned)[47] <- "FRUIT"
names(brfss_2017_cleaned)[48] <- "FRUIT_JUICE"
names(brfss_2017_cleaned)[49] <- "GREEN_VEG"
names(brfss_2017_cleaned)[50] <- "FRENCHF"
names(brfss_2017_cleaned)[51] <- "POTATO"
names(brfss_2017_cleaned)[52] <- "OTHER_VEG"
names(brfss_2017_cleaned)[53] <- "EXERCISE_ANY"
names(brfss_2017_cleaned)[56] <- "FLUSHOT"
names(brfss_2017_cleaned)[57] <- "PNEUMONIA_VAC"
names(brfss_2017_cleaned)[58] <- "HIVTEST"
names(brfss_2017_cleaned)[59] <- "HIVRISK"
names(brfss_2017_cleaned)[63] <- "STSTR"
names(brfss_2017_cleaned)[64] <- "STRATUM_WGHT"
names(brfss_2017_cleaned)[65] <- "RAWRAKE"
names(brfss_2017_cleaned)[66] <- "WT2RAKE"
names(brfss_2017_cleaned)[67] <- "IMPRACE"
names(brfss_2017_cleaned)[68] <- "DUALUSE"
names(brfss_2017_cleaned)[69] <- "DUALCOR"
names(brfss_2017_cleaned)[70] <- "LLCPWT2"
names(brfss_2017_cleaned)[71] <- "LLCPWT"
names(brfss_2017_cleaned)[72] <- "RFHLTH"
names(brfss_2017_cleaned)[73] <- "PHYS_COMP"
names(brfss_2017_cleaned)[74] <- "MENT_COMP"
names(brfss_2017_cleaned)[75] <- "HCVU651"
names(brfss_2017_cleaned)[76] <- "RFHYPES"
names(brfss_2017_cleaned)[77] <- "CHOLCH1"
names(brfss_2017_cleaned)[78] <- "HICHOL_COMP"
names(brfss_2017_cleaned)[79] <- "MI_CHD"
names(brfss_2017_cleaned)[80] <- "ASTHMA_LIFE"
names(brfss_2017_cleaned)[81] <- "ASTHMA_NOW_COMP"
names(brfss_2017_cleaned)[82] <- "ASTHMA_COMP"
names(brfss_2017_cleaned)[83] <- "ARTHRITIS_DIAGNOS"
names(brfss_2017_cleaned)[84] <- "LIMIT_ACTIVITY"
names(brfss_2017_cleaned)[85] <- "LIMIT_WORK"
names(brfss_2017_cleaned)[86] <- "LIMIT_SOCIAL"
names(brfss_2017_cleaned)[87] <- "PRACE_COMP"
names(brfss_2017_cleaned)[88] <- "MULTI_RACE"
names(brfss_2017_cleaned)[89] <- "HISPANIC_COMP"
names(brfss_2017_cleaned)[90] <- "RACE_COMP"
names(brfss_2017_cleaned)[91] <- "RACEG21"
names(brfss_2017_cleaned)[92] <- "RACEGR3"
names(brfss_2017_cleaned)[93] <- "RACE_G1"
names(brfss_2017_cleaned)[94] <- "AGEG5YR"
names(brfss_2017_cleaned)[95] <- "AGE65YR"
names(brfss_2017_cleaned)[96] <- "AGE80"
names(brfss_2017_cleaned)[97] <- "AGE_G"
names(brfss_2017_cleaned)[98] <- "HT_COMP_IN"
names(brfss_2017_cleaned)[99] <- "HT_COMP_M"
names(brfss_2017_cleaned)[100] <- "WT_COMP_KILO"
names(brfss_2017_cleaned)[101] <- "BMI"
names(brfss_2017_cleaned)[102] <- "BMI_CAT"
names(brfss_2017_cleaned)[103] <- "BMI_OVER"
names(brfss_2017_cleaned)[104] <- "CHILD_COMP"
names(brfss_2017_cleaned)[105] <- "EDUCA_COMP"
names(brfss_2017_cleaned)[106] <- "INCOME_COMP"
names(brfss_2017_cleaned)[107] <- "SMOKER_COMP"
names(brfss_2017_cleaned)[108] <- "SMOKENOW_COMP"
names(brfss_2017_cleaned)[109] <- "ECIG_COMP"
names(brfss_2017_cleaned)[110] <- "ECIG_NOW_COMP"
names(brfss_2017_cleaned)[111] <- "DRINKANY_30"
names(brfss_2017_cleaned)[112] <- "DROCDY3"
names(brfss_2017_cleaned)[113] <- "BINGE_COMP"
names(brfss_2017_cleaned)[114] <- "WEEKLY_DRINK"
names(brfss_2017_cleaned)[115] <- "HEAVY_DRINK"
names(brfss_2017_cleaned)[116] <- "FRUITJUICE_COMP"
names(brfss_2017_cleaned)[117] <- "FRUITJUICE_COMP2"
names(brfss_2017_cleaned)[118] <- "GREEN_VEG_COMP"
names(brfss_2017_cleaned)[119] <- "FRENCHF_COMP"
names(brfss_2017_cleaned)[120] <- "POTATO_COMP"
names(brfss_2017_cleaned)[121] <- "OTHER_VEG_COMP"
names(brfss_2017_cleaned)[122] <- "MISSING_FRUIT"
names(brfss_2017_cleaned)[123] <- "MISSING_VEG"
names(brfss_2017_cleaned)[124] <- "MISS_ANY_FRUIT"
names(brfss_2017_cleaned)[125] <- "MISS_ANY_VEG"
names(brfss_2017_cleaned)[126] <- "TOTAL_FRUIT"
names(brfss_2017_cleaned)[127] <- "TOTAL_VEG"
names(brfss_2017_cleaned)[128] <- "ATLEAST1_FRUIT"
names(brfss_2017_cleaned)[129] <- "ATLEAST1_VEG"
names(brfss_2017_cleaned)[130] <- "MORE16_FRUIT"
names(brfss_2017_cleaned)[131] <- "MORE23_VEG"
names(brfss_2017_cleaned)[132] <- "FRUIT_EXCLUDE"
names(brfss_2017_cleaned)[133] <- "VEG_EXCLUDE"
names(brfss_2017_cleaned)[134] <- "LEISURE_COMP"
names(brfss_2017_cleaned)[135] <- "MAX_OXYGEN"
names(brfss_2017_cleaned)[136] <- "FUNC_CAPACITY"
names(brfss_2017_cleaned)[137] <- "STRENGTH_FREQ"
names(brfss_2017_cleaned)[138] <- "MISSING_PHYS"
names(brfss_2017_cleaned)[139] <- "PHYS_CATEGORY"
names(brfss_2017_cleaned)[140] <- "PHYS_IDX"
names(brfss_2017_cleaned)[141] <- "150_PHYS_COMP"
names(brfss_2017_cleaned)[142] <- "300_PHYS_COMP"
names(brfss_2017_cleaned)[143] <- "300_PHYS2_COMP"
names(brfss_2017_cleaned)[144] <- "MUSCLE_RECOMMEND"
names(brfss_2017_cleaned)[145] <- "AEROBIC_STRENGTH"
names(brfss_2017_cleaned)[146] <- "AEROBIC_STRENGTH2"
names(brfss_2017_cleaned)[147] <- "SEATBELT_COMP"
names(brfss_2017_cleaned)[148] <- "SEATBELT_COMP2"
names(brfss_2017_cleaned)[149] <- "AIDSTEST_COMP"

head(brfss_2017_cleaned)
```

Missing Data Summary Table

```{r}
missing_summary <- brfss_2017_cleaned %>%
  miss_var_summary()

print(missing_summary)
```

Value Imputation

```{r}
# Replace NA values with O for entire data frame
brfss_2017_cleaned[is.na(brfss_2017_cleaned)] <- 0

# Check if any NA values remain
colSums(is.na(brfss_2017_cleaned))

# Create VapeStatus based on ECIG_COMP
df <- brfss_2017_cleaned %>%
    filter(!is.na(ECIG_COMP)) %>%
    mutate(VapeStatus = factor(ECIG_COMP, 
                               levels = c(1, 2, 3, 4, 9),
                               labels = c("Current User", "Current User", "Former User", "Never Used", "No Response")))
    
# Re-code STATE column
df$STATE_ABBR <- recode(df$STATE,
                       '1' = 'AL',
                       '2' = 'AK',
                       '4' = 'AZ',
                       '5' = 'AR',
                       '6' = 'CA',
                       '8' = 'CO',
                       '9' = 'CT',
                       '10' = 'DE',
                       '11' = 'DC',
                       '12' = 'FL',
                       '13' = 'GA',
                       '15' = 'HI',
                       '16' = 'ID',
                       '17' = 'IL',
                       '18' = 'IN',
                       '19' = 'IA',
                       '20' = 'KS',
                       '21' = 'KY',
                       '22' = 'LA',
                       '23' = 'ME',
                       '24' = 'MD',
                       '25' = 'MA',
                       '26' = 'MI',
                       '27' = 'MN',
                       '28' = 'MS',
                       '29' = 'MO',
                       '30' = 'MT',
                       '31' = 'NE',
                       '32' = 'NV',
                       '33' = 'NH',
                       '34' = 'NJ',
                       '35' = 'NM',
                       '36' = 'NY',
                       '37' = 'NC',
                       '38' = 'ND',
                       '39' = 'OH',
                       '40' = 'OK',
                       '41' = 'OR',
                       '42' = 'PA',
                       '44' = 'RI',
                       '45' = 'SC',
                       '46' = 'SD',
                       '47' = 'TN',
                       '48' = 'TX',
                       '49' = 'UT',
                       '50' = 'VT',
                       '51' = 'VA',
                       '53' = 'WA',
                       '54' = 'WV',
                       '55' = 'WI',
                       '56' = 'WY',
                       '66' = 'GU',
                       '72' = 'PR')
```

Train-Test Split Ratio

```{r}
# Filter for only asthmatics
df <- df %>%
  filter(ASTHMA == 1)

# Linear regression
set.seed(123)
x <- data.matrix(df[, c('AGE80', 'BMI', 'SMOKER_COMP', 'STATE', 'ADDEPEV', 'ALCDAY5', 'CHCCOPD', 'CHECKUP', 'EDUCA', 'GENHLTH', 'HLTHPLN', 'INCOME', 'SEX')])
y <- df$ECIG_COMP
data <- data.frame(x, y)

# Fit the linear regression model
model <- lm(y ~ x, data = data)
plot(model)

# Calculate model parameters
result <- model_parameters(model)
result

plot(result)

# Calculate the regression coefficients (slope and intercept)
coefficients <- coef(model)
slope <- coefficients[2] # Slope coefficient
intercept <- coefficients[1] # Intercept

# Make predictions
predictions <- predict(model, newdata = data.frame(x))

# Calculate the squared error loss
actual_values <- data$y
squared_error_loss <- sum((actual_values - predictions)^2)

# Calculate the Mean Squared Error (MSE)
mse <- mean((actual_values - predictions)^2)
mse

# Perform k-fold cross-validation to find optimal lambda value
cv_model <- cv.glmnet(x, y, alpha = 1)

# Produce plot of test MSE by lambda value
plot(cv_model)

# Optimal splitting ratio
random_values <- rnorm(n = 62411, mean = 0, sd = 1)
optimal_ratio <- splitratio(random_values, y, method = "regression")
optimal_ratio

# Create training & test sets
set.seed(123)
split <- sort(sample(nrow(df), nrow(df) * 0.6)) 
train <- df[split,]
test <- df[-split,]


calcSplitRatio <- function(p = NA, df) {
  ## p  = the number of parameters. by default, if none are provided, the number of columns (predictors) in the dataset are used
  ## df = the dataframe that will be used for the analysis
  
  ## If the number of parameters isn't supplied, set it to the number of features minus 1 for the target
  if(is.na(p)) {
    p <- ncol(df) -1   ## COMMENT HERE
  }
  
  ## Calculate the ideal number of testing set
  test_N <- (1/sqrt(p))*nrow(df)
  ## Turn that into a testing proportion
  test_prop <- round((1/sqrt(p))*nrow(df)/nrow(df), 2)
  ## And find the training proportion
  train_prop <- 1-test_prop
  
  ## Tell us the results!
  print(paste0("The ideal split ratio is ", train_prop, ":", test_prop, " training:testing"))
  
  ## Return the size of the training set
  return(train_prop)
}
```

Encoding

```{r}
# No categorical variables to encode - ECIG_COMP is the encoding of VapeStatus & STATE is the encoding of STATE_ABBR
# Remove categorical variables from dataset
df <- df %>%
    select(-c(150:151))
```

Arithmetic Transformations

```{r}
# Logarithmic Transformation
df$log_x <- log(df$ECIG_COMP)

# Square Root Transformation
df$sqrt_x <- sqrt(df$ECIG_COMP)
```

Normalize and/or Scale

```{r}
# Applying preProcess for Min-Max normalization
preprocessed_data <- preProcess(df[, -c(1:4, 18, 80)], method = "range") # Remove first 4 columns since they are unique variables; Remove ASTHMA & ASTHMA_LIFE since they have no variation
transformed_data <- predict(preprocessed_data, df)

print(transformed_data)
```


# Feature Engineering

```{r}
# Remove character columns
numerical_data <- transformed_data %>%
    select(-c(3))

# Select only predictor & target variables
numerical_data <- numerical_data %>% 
    select(c('STATE', 'GENHLTH', 'HLTHPLN', 'CHECKUP', 'CHCCOPD', 'ADDEPEV', 'SEX', 'EDUCA', 'INCOME', 'ALCDAY5', 'AGE80', 'BMI', 'SMOKER_COMP', 'ECIG_COMP'))

# PCA
pca <- princomp(numerical_data)
summary(pca)

# Use the loading of each principal component to understand the relation to each column
pca$loadings[, 1:2]

# Visualize loading
fviz_eig(pca, addlabels = TRUE)
```

